### 2012-11-16 doc/progress_plan/appendix.tex
Finished compiling appendicies

\appendix
\section{User Interface Designs}
\label{sec:ui_designs}

\subsection{ Partridge Query Interface}

Figure \ref{fig:ui_mockup} shows the user interface that will be utilised to
search and query Partridge's corpus.

The query form shows two main sections: Filters and Keywords.

Filters can be used to show a large set of papers where the user is unsure of
what to search for. Users will be able to look for papers by type and topic as
discussed above. The paper result filter has not been included in this diagram.
However, it is expected that a drop down menu for result type would be included
in the filters section of the form.

Keywords allows the user to look for a set of keywords that is specifically
within a CoreSC concept for a paper. For example, the reader may want to find
experiments that use a server farm to do lots of calculations. Therefore, they
would enter ``Server Farm" as their keywords and chose ``Method" from the paper
section. The user then adds this to the set of keyword queries in the table
below the text field with the add button.

Once a user has configured both filters and keywords to their preferences, they
click the search form to run the search itself.

Other notable features on this mockup include a list of most recent searches.
This helps the user to understand how to use Partridge and gives them some
suggestions for what they might search for. Similar listings are provided for
the most popular papers in the Partridge database and the most recent papers
added to the system.

\begin{figure}[!ht]
\includegraphics[width=\textwidth]{images/mockup_1.png}
\caption{User interface design for querying Partridge}
\label{fig:ui_mockup}
\end{figure}

\section{System Process Diagrams}
\label{sec:system_diagrams}

This section documents the system process diagrams that have been produced. The
first completed diagram shows how a paper will be added to Partridge and the
actions that will be carried out upon it.

\subsection{Adding a Paper to Partridge}

\begin{figure}[!ht]
\includegraphics[width=0.75\textwidth]{images/PaperAddedProcess.png}
\caption{The process of adding a new paper to Partridge}
\end{figure}

The system starts off by converting the document to XML from PDF if necessary
through the use of the PDFX tool provided by Manchester University. A sentence
splitting tool is then used to parse the resulting document and separate the
sentences.

Once the sentences have been split, Maria's SAPIENTA tool is used to annotate
each sentence to determine which of the core scientific concepts it covers.
This information is stored back in the XML file with the document.

After this, feature extraction is carried out upon the document to find useful
features for the following classification tasks:

\begin{enumerate}
\item Paper topic/subject -i.e. is the paper a biology paper or a computer science paper?
\item Type of paper - i.e. is the paper a case study, an experiment or a literature review?
\item Paper result - i.e. did the paper have a positive, negative or inconclusive result?
\end{enumerate}

These classifiers are then run (they could potentially be run in parallel if
processing power is available) to determine the paper's class for each
classifier respectively. The data gathered along with any paper metadata
captured such as title, author, date, institute etc are then stored in the
database.

\section{Email for CRFSuite}
\label{sec:crfemail}
\input{crfletter}

\section{Project Timeline}
\label{sec:timeline}

This table shows the projected timeline for the Partridge project. These
calculations were carried out under the assumption of 2 full days of work per
week on the project and one iteration equating to one month (28 days). There
will be 7 iterations in total and after each, a new version of Partridge will
be made available (except for iteration zero which is a research only
iteration).

\begin{figure}[htp] 
 \vspace{ -1cm }
 \centering{\includegraphics[scale=0.60,page=1]{../timeline.pdf}}

\end{figure}

\begin{figure}[!htp] 
 \centering{\includegraphics[scale=0.60,page=2]{../timeline.pdf}}
 \vspace{-2cm}
\centering{\includegraphics[scale=0.60,page=3]{../timeline.pdf}}
\end{figure}

\pagebreak
\section{Project Wiki}
\label{sec:wiki}

All pages from the Wiki are shown below with the exceptions of those pages
which have already been mostly featured in the document (UI Mockups and
Progress Diagrams and the project timetable are not shown in this section).

\subsection{ Notes 03/10/2012 }
\input{wiki/Notes-03-10-2012}

\subsection{ Notes 10/10/2012 }
\input{wiki/Notes-10-10-2012}

\subsection{ Notes 19/10/2012 }
\input{wiki/Notes-19-10-2012}

\subsection{ Notes 24/10/2012 }
\input{wiki/Notes-24-10-2012}

\subsection{ Notes 31/10/2012 }
\input{wiki/Notes-31-10-2012}

\subsection{ Notes 07/11/2012 }
\input{wiki/Notes-07-11-2012}

\pagebreak
\bibliographystyle{IEEEannot}
\bibliography{report}