### 2012-11-16 doc/progress_plan/wiki/Notes-03-10-2012.tex

\subsection{Meeting Notes from 03/10/2012}

\subsubsection{General topics}

\begin{itemize}
\item
  Talked about using research papers instead of books :-
\item
  copyright issues - mainly resolved because of Gutenburg project
\item
  processing issues - too much information to process quickly and
  realistically
\item
  Maria has some software for identifying sections in papers - could be
  used to help locate specific features
\item
  There are quite a few sources for papers that could be used for the
  project:
\item
  http://arxiv.org/
\item
  http://plosone.org/
\item
  Maria's papers are in XML format
\item
  Formatting may or may not be an issue - PDFs are a nightmare to parse
\end{itemize}

\subsubsection{Things to consider}

\begin{itemize}
\item
  What sort of recommendations for different sources might there be and
  how would these recommendations be extracted?
\item
  Feature engineering - what sorts of features might there be to pick
  out in the papers?
\item
  I need to have a read through the sentiment analysis paper that Maria
  has been working on
\item
  I need to have a play with the software for identifying sections in
  papers.
\end{itemize}

\subsubsection{Misc concerns}

\begin{itemize}
\item
  Decided that our weekly meeting should be Wednesdays at 3pm
\item
  Investigate setting up source control and a wiki (that's been done as
  you can tell)
\end{itemize}