### 2012-11-16 doc/progress_plan/wiki/Notes-31-10-2012.tex

\subsection{James' Meeting Notes}

\subsubsection{To Think about}

\begin{itemize}
\item
  Discussed imminent deadline for project progress report - 19/11/2012
\item
  Decided that having a report finished by 14/11/2012 would be a good
  report to give Amanda and Maria a chance to check over the report
  before handin.
\item
  Maria explained that it is possible to send a batch of papers to
  Sapienta rather than one at a time. She is going to provide
  information on how this can be achieved at some point during the week.
\item
  I should describe my working processes and why they are helpful.
\end{itemize}

\subsubsection{Todo}

\begin{itemize}
\item
  We discussed the merits of using the pdfx converter instead of my
  script - might have plug-and-play compatibility with Sapienta for a
  start - need to verify this.
\item
  I need to send my error message the CRFSuite compilation to Maria so
  that she can contact the maintainer.
\item
  I need to expand on the system outline providing more specific goals
  and timeframes - this will become my development plan.
\item
  I need to look at textpresso and see if it will help me with my
  comparison/evaluation
\end{itemize}