### 2012-11-18 doc/progress_plan/wiki/Notes-07-11-2012.tex

\subsection{Notes from 7/11/2012}

\subsubsection{Discussed}

\begin{itemize}
\item
  What I'll need to have for my demo:

  \begin{itemize}
  \item
    Simple web interface
  \item
    Few basic classifiers - probably result of paper positive/negative
    and maybe paper type - proportions of coreSC concepts.
  \item
    Could show off PDF to XML converter script
  \end{itemize}
\end{itemize}

\subsubsection{To Consider}

\begin{itemize}
\item
  Is the disparity between bio paper CoreSC classification and other
  sorts of paper to do with Sapienta's model or does it demonstrate a
  feature - different subjects may have different proportions of each
  concept.
\item
  If I'm using distribution of concepts as an aggregate feature, I need
  to make sure the system does not fall over if a concept is missing.
\item
  Front end interface:
\item
  Need to find a new word for `section' Scientific Concept may be a bit
  too `scary', is there a better term for this?
\item
  If I allow users to upload papers, they must sign a digital disclaimer
  to say that they have the right to do so
\item
  There should be a ``report copyrighted material'' system
\end{itemize}

\subsubsection{To Do}

\begin{itemize}
\item
  Biggest task this week: Progress Report.

  \begin{itemize}
  \item
    Draft due 14/11/2012 for discussion on 15/11/2012
  \item
    Final deadline 19/11/2012
  \item
    Could provide wiki content as an appendix to the report
  \item
    Come up with a project timeline - actually consider dates
  \item
    Start with the web interface and work upwards.
  \end{itemize}
\item
  Send Maria the missing header file from crfsuite
\item
  Link system outline from wiki index
\item
  Maria said she'd look for a topic detection paper for me to peruse
\item
  I need to have another look at relevant features - struggling with
  this
\item
  Make a wiki list of features that could be used for classification
\end{itemize}