### 2012-11-18 doc/progress_plan/wiki/Notes-19-10-2012.tex

\subsection{Notes from Meeting 18/10/2012}

\subsubsection{General Comments}

\begin{itemize}
\item
  I need to decide on a domain for the project. It may be that papers on
  a varied selection of topics would be a good start
\item
  Decided on an agile development cycle with a basic program that is
  improved in `iterations'.
\item
  As part of testing the software, it may be possible to get other final
  year students to trial the engine and use it to make recommendations
  for their reading.
\item
  I need to investigate ways of analysing papers for syntactic patterns.
  This may be indicative of:

  \begin{itemize}
  \item
    Different authors' writing styles
  \item
    Different types of paper -i.e.~psychology vs physics
  \end{itemize}
\end{itemize}

\subsubsection{TODO}

\begin{itemize}
\item
  Revisit PDF to XML conversion as conversion from PDF would make life a
  lot easier
\item
  Maria mentioned that she may have access to an HTML to XML program
  which would be good to look at.
\item
  Check out python Jerboa to see if it would be useful for the project.
\end{itemize}