### 2013-02-26 doc/final_report/report.tex
started work on final report

% Final Dissertation Report

%The document is a report
\documentclass[11pt,a4paper]{article}

%define horizontal rule
\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}

\usepackage{fullpage}
\usepackage{pdfpages}
%use the listings package
\usepackage{listings}
%use the English language
\usepackage[english]{babel}
%use graphics
\usepackage{graphicx}
%use wrap figures
\usepackage{wrapfig}
%geometry stuffs
\usepackage{lscape}
%use natbib bibliography package
%\usepackage[numbers]{natbib}
%use harvard bibliography package
%\usepackage{harvard}	
%use captions
\usepackage{caption}
%use multi-row tables
\usepackage{multirow}
\usepackage{url}
\usepackage{subcaption}

\usepackage[nottoc,numbib]{tocbibind}

\begin{document}

%include the title page
\input{revision}
\input{title}

%some definitions for paragraph layout stuff
\setlength{\parindent}{0pt}
\setlength{\parskip}{1.5ex plus 0.5ex minus 0.2ex}

%TODO: STATEMENT OF ORIGINALITY

%TODO: ACKNOWLEDGEMENTS

%TODO: ABSTRACT

\tableofcontents

\pagebreak

\section{Introduction}

\pagebreak
\bibliographystyle{IEEEannot}
\bibliography{report}

\end{document}