### 2013-03-06 doc/final_report/report.tex
Made some changes to the intro

% Final Dissertation Report

%The document is a report
\documentclass[11pt,a4paper,oneside]{book}

%define horizontal rule
\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}

\usepackage{fullpage}
\usepackage{pdfpages}
%use the listings package
\usepackage{listings}
%use the English language
\usepackage[english]{babel}
%use graphics
\usepackage{graphicx}
%use wrap figures
\usepackage{wrapfig}
%geometry stuffs
\usepackage{lscape}
%use natbib bibliography package
%\usepackage[numbers]{natbib}
%use harvard bibliography package
%\usepackage{harvard}	
%use captions
\usepackage{caption}
%use multi-row tables
\usepackage{multirow}
\usepackage{url}
\usepackage{subcaption}

\usepackage[nottoc,numbib]{tocbibind}

\usepackage{fancyhdr}
\setlength{\headheight}{15pt}

\pagestyle{fancy}

\begin{document}

%include the title page
\input{revision}
\input{title}

%some definitions for paragraph layout stuff
\setlength{\parindent}{0pt}
\setlength{\parskip}{1.5ex plus 0.5ex minus 0.2ex}

\thispagestyle{plain}

\begin{center}
\textsc{\large Statement of Originality}
\end{center}

This submission is my own work, except where clearly indicated.

I understand that there are severe penalties for plagiarism and other unfair
practice, which can lead to loss of marks or even the withholding of a degree.
I have read the sections on unfair practice in the Students’ Examinations
Handbook and the relevant sections of the current Student Handbook of the
Department of Computer Science.

I understand and agree to abide by the University’s regulations governing these
issues.\\\\

........................................

\pagebreak
\thispagestyle{plain}

\begin{center}
\textsc{\large Acknowledgements}
\end{center}

I would like to take this opportunity to say thank you to both of my
supervisors, Amanda Clare and Maria Liakata, for giving up so much of their
time to support me in designing and developing this project.

Thanks to the users who tested Partridge once it was made generally available
and for their feedback in the form of bug reports and suggestions.

Finally, I am eternally grateful for the support of my family and friends who
have put up with my gripes and moans throughout the duration of this project
and without whom, I couldn't have achieved such an ambitious project.

\pagebreak
\thispagestyle{plain}

\begin{center}
\textsc{\large Abstract}
\end{center}

When faced with the great number of scientific papers made general available on
the internet, it is often very difficult for researchers to find relevant
scientific literature. This project aimed to build an information retrieval and
recommendation tool that uses Natural Language Processing and Machine Learning
techniques to present the most relevant scientific papers to a user based upon
their search criteria and preferences.

A preprocessing system system that exploits existing tools such as SAPIENTA and
PDFX was used to analyse a large sample of research papers, labelling the
presence of core scientific concepts (CoreSCs) within each paper and extracting
metadata. Decision trees were also trained to classify papers into categories
such as `Research Paper', `Review Paper' and `Case Study' based upon the
relative proportions of core scientific concepts in the papers.

A web interface was built to facilitate user search of papers based upon the
contents of specific CoreSCs. Filtering of papers using their type as
determined by the preprocessor was also built in and a page allowing authors and
rightsholders to submit their own papers for analysis was created to allow
the scientific community to extend the utility of the system.

\tableofcontents

\pagebreak

\mainmatter

\chapter{Introduction}
\input{introduction}

\chapter{ Development Methodology }
\input{methodology}

\chapter{Design}
\input{design}

\chapter{Implementation}
\input{implementation}

\chapter{Testing}
\input{testing}

\chapter{Evaluation}
\input{evaluation}

\pagebreak
\bibliographystyle{IEEEannot}
\bibliography{report}

\appendix
\input{appendix}

\end{document}