### 2013-03-18 doc/final_report/design.tex
Finished first draft of testing chapter and rearranged contents

%%-----------------------------------------------
%
% Include for design chapter of dissertation
%
%%----------------------------------------------

\section{ System Architecture}

It was decided from the outset that Partridge would be a web-based tool.
Web-based systems can generally be run on any computer with access to the
internet and a modern web browser. This also means that there is no need for
the end user to install or configure extra software on their computers, making
Partridge accessible to non-technical users and those who have aggressive
software restrictions on their computers, such as GPs and users of public
computers in libraries or on academic sites.

\section{ Subsystem Design }

\subsection{ Preprocessor }

\subsection{ Web Backend }

\subsection{ Web Frontend }