### 2013-03-27 doc/final_report/crfletter.tex
Finished implementation section on clustering and paper type classification

\begin{verbatim}
Hi All,

I'm trying to build CRFSuite's python extension 
(http://www.chokkan.org/software/crfsuite/) on an Arch Linux 64-bit 
environment using Swig 2.0.8 (PCRE enabled).

When I run SWIG, I get the following output:

<<OUTPUT OMITTED FOR BREVITY>>

Full output available at http://sourceforge.net/mailarchive/message.php?msg_id=30024209

As far as I can see, all of these problems stem from this declaration in 
the SWIG input file (http://pastebin.com/9JipJ1C1):
%template(ItemSequence) std::vector<CRFSuite::Item>;

I've had a look around on google and on this mailing list and the only 
other "cannot copy typemap" errors I've encountered have been where 
people have excluded the std namespace in favour of a 'using' statement. 
As you can see, this example uses absolute class names so that isn't the 
issue here. I haven't had any luck contacting the software maintainer 
for CRFSuite yet either.

As you might expect, if I continue to compile the extension without 
heeding these errors, when I try to make use of the ItemSequence object 
I get the following python error:
TypeError: in method 'ItemSequence_append', argument 2 of type 
'std::vector< CRFSuite::Item >::value_type const &'

I'm going to keep trying to sort this out and if I find the solution 
independantly, I'll make sure to post a follow-up email. However, if 
anyone else has any ideas about what might be happening here, please let 
me know.

Thanks,

James Ravenscroft
AI & Robotics Undergraduate
Aberystwyth University
\end{verbatim}