### 2013-04-15 doc/final_report/report.tex

% Final Dissertation Report

%The document is a report
\documentclass[11pt,a4paper,oneside]{book}

%define horizontal rule
\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}

\usepackage{fullpage}
\usepackage{pdfpages}
%use the listings package
\usepackage{listings}
%use the English language
\usepackage[english]{babel}
%use graphics
\usepackage{graphicx}
%use wrap figures
\usepackage{wrapfig}
%geometry stuffs
\usepackage{lscape}
%use captions
\usepackage{caption}
%use multi-row tables
\usepackage{multirow}
\usepackage{url}
\usepackage{breakurl}
\usepackage{subcaption}
\usepackage{rotating}
\usepackage{pdflscape}

\usepackage[nottoc,numbib]{tocbibind}

\usepackage{fancyhdr}
\setlength{\headheight}{15pt}

\pagestyle{fancy}

\begin{document}

%include the title page
\input{revision}
\input{title}

%some definitions for paragraph layout stuff
\setlength{\parindent}{0pt}
\setlength{\parskip}{1.5ex plus 0.5ex minus 0.2ex}

\thispagestyle{plain}

\begin{center}
\textsc{\large Statement of Originality}
\end{center}

This submission is my own work, except where clearly indicated.

I understand that there are severe penalties for plagiarism and other unfair
practice, which can lead to loss of marks or even the withholding of a degree.
I have read the sections on unfair practice in the Students’ Examinations
Handbook and the relevant sections of the current Student Handbook of the
Department of Computer Science.

I understand and agree to abide by the University’s regulations governing these
issues.\\\\

........................................

\pagebreak
\thispagestyle{plain}

\begin{center}
\textsc{\large Acknowledgements}
\end{center}

I would like to take this opportunity to thank both of my
supervisors, Amanda Clare and Maria Liakata, for giving up so much of their
time to support me in designing and developing this project.

Thanks to the users who tested Partridge once it was made generally available
and for their feedback in the form of bug reports and suggestions.

Finally, I am eternally grateful for the support of my family and friends who
have put up with my gripes and moans throughout the duration of this project
and, without whom, I couldn't have achieved such an ambitious project.

\pagebreak
\thispagestyle{plain}

\begin{center}
\textsc{\large Abstract}
\end{center}

When faced with the great number of scientific papers made general available on
the internet, it is often very difficult for researchers to find relevant
scientific literature. This project aimed to build an information retrieval and
recommendation tool that uses Natural Language Processing and Machine Learning
techniques to present the most relevant scientific papers to a user based upon
their search criteria and preferences.

A preprocessing system that exploits existing Natural Language Processing tools
was used to analyse a large sample of research papers, labelling the presence
of conceptual zones within each paper and extracting metadata.  Decision trees
were also trained to classify papers into well established type categories
based upon the relative proportions of the types of conceptual zones
present within the papers.

A web interface was built to facilitate fine grain search, making use of the
zone information captured during preprocessing. Filtering of papers using their
type as determined by the preprocessor was also built in. Individual papers
returned during searches are presented in their own profile page that provides
charts and extra metadata to help researchers to understand the paper's
content. A page allowing authors and rightsholders to submit their own papers
for analysis was also created to allow the scientific community to extend the
impact and utility of the system.

The machine learning systems used within the project have been evaluated and
found to be accurate to 87.9\% when classifying a paper's type. The possibility
of extending the machine learning capabilities of the system to categorise
papers by subject domain and outcome of experimental results is discussed as
well as the potential for further improvements to the user interface.

A working version of the software has been made generally available at
\url{http://farnsworth.papro.org.uk} and can be explored and used by the
general public. It is hoped that over the coming months, the system will grow
to become a far reaching, widely used tool for academic research.

\tableofcontents

\pagebreak

\mainmatter

\chapter{Introduction}
\input{introduction}

\chapter{ Development Methodology }
\input{methodology}

\chapter{Design}
\input{design}

\chapter{Implementation}
\input{implementation}

\chapter{Testing}
\label{chapter:testing}
\input{testing}

\chapter{Critical Evaluation}
\input{evaluation}

\pagebreak
\bibliographystyle{IEEEannot}
\bibliography{report}

\appendix
\input{appendix}

\end{document}