### 2013-05-24 doc/bcspaper/article.tex
Maria edits all sections

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a contributed volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass{svmult}
%
%% choose options for [] as required from the list
%% in the Reference Guide
%
\usepackage{mathptmx}       % selects Times Roman as basic font
\usepackage{helvet}         % selects Helvetica as sans-serif font
\usepackage{courier}        % selects Courier as typewriter font
\usepackage{type1cm}        % activate if the above 3 fonts are
                             % not available on your system
%
\usepackage{makeidx}         % allows index generation
\usepackage{graphicx}        % standard LaTeX graphics tool
%                             % when including figure files
\usepackage{multicol}        % used for the two-column index
\usepackage[bottom]{footmisc}% places footnotes at page bottom

%\usepackage{subcaption}

%\captionsetup{compatibility=false}

%
%% see the list of further useful packages
%% in the Reference Guide
%
%\makeindex             % used for the subject index
%                       % please use the style svind.ist with
%                       % your makeindex program
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\begin{document}

\title*{Partridge: An effective system for the automatic classification of the types of academic papers}
\titlerunning{Partridge: classifying academic papers}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{James Ravenscroft \and Maria Liakata \and Amanda Clare}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{James Ravenscroft, Amanda Clare \at Department of Computer Science, Aberystwyth University, Aberystwyth, SY23 3DB, UK  \email{jrr9@aber.ac.uk}
\and Maria Liakata \at Department of Computer Science, University of Warwick \email{M.Liakata@warwick.ac.uk}}

%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle

\abstract{Partridge is a system that enables intelligent search for
  academic papers by allowing users to query terms within
  sentences designating a particular core scientific concept (e.g. Hypothesis, Result, etc). 
  The system also automatically classifies papers according to
  article types (e.g. Review, Case Study). Here we focus on the latter
  aspect of the system. For each paper, Partridge automatically
  extracts the full paper content from PDF files, converts it to XML,
  determines sentence boundaries, automatically labels the sentences
  with core scientific concepts, and then uses a random forest model to
  classify the paper type. We show that the type of a paper can be
  reliably predicted by a model which analyses the distribution of
  core scientific concepts within the sentences of the paper. We
  discuss the appropriateness of many of the existing paper types used
  by major journals, and their corresponding distributions. Partridge
  is online and available for use, includes a browser-friendly
  bookmarklet for new paper submission, and demonstrates a range of
  possibilities for more intelligent search in the scientific
  literature. The Partidge instance can be used at
  \url{http://farnsworth.papro.org.uk}, and further information about
  the project can be found at \url{http://papro.org.uk}}

\section{Introduction} \label{sec:1} Since the advent of the `Digital Age' and
the ability of computers to copy and reproduce information for a negligible
cost, the amount of information available to researchers has been increasing
drastically. As available information increases, relevant material becomes
progressively more difficult to find manually and the need for an automated
information retrieval tool more apparent. There are already a large number of
information retrieval and recommendation systems for scientific publications.
Many of these systems, such as AGRICOLA (\url{http://agricola.nal.usda.gov/}),
the Cochrane Library(\url{http://www.thecochranelibrary.com/}) and Textpresso
(\url{http://www.textpresso.org/}) index only publications from predefined
journals or topics (for the above examples, Agriculture, Biology and
Bioinformatics respectively).  Unfortunately, these domain specific indexing
systems usually only contain a small subset of papers, excluding potentially
crucial literature because it does not quite fit into the subject domain.

The value of these systems to their users is often restricted by the small
proportion of available literature that they index, forcing researchers to use
multiple, domain specific, search engines for their queries.  In contrast,
there are also a number of interdisciplinary indexing systems and online
journals such as arXiv( \url{http://arxiv.org/}),
PloSOne(\url{http://plosone.org/}), and JSTOR (\url{http://www.jstor.org/}),
that try to incorporate wide ranges of papers from as many disciplines as
possible. The traits of these systems often complement those of their
domain-specific counterparts; they provide a comprehensive collection of
literature but insufficient filtering and indexing capabilities.  One of the
most publicised and well known interdisciplinery scientific literature search
systems is Google Scholar (\url{http://scholar.google.com}). Google offers
advanced query options specific to Scholar that allow searching by author, year
and for words that occur only in the document title.

However, the document title is just one of the crucial parts of a scientific
paper's structure.  Liakata \emph{et al.} (2012) describe a system for
automatically processing and classifying sentences in a research paper
according to the core scientific concept (CoreSC) that they
describe\cite{Liakata2012}.  There are 11 CoreSCs, including {\em Hypothesis},
{\em Goal}, {\it Background}, {\em Method}, {\em Result} and {\em Conclusion}.
CoreSC labels can be allocated to all sentences in a scientific paper in order
to identify which scientific concept each sentence encapsulates.  SAPIENTA
(\url{http://www.sapientaproject.com}) is a publicly available machine learning
application which can automatically annotate all sentences in a scientific
paper with their CoreSC labels. It was trained using a corpus of physical
chemistry and biochemistry research papers whose sentences were manually
annotated using the CoreSC\cite{LIAKATA10.644} scheme.  An intelligent academic
information retrieval system can use this data to provide better filtering and
search capabilities for researchers.  Partridge implements such context-aware
keyword search, by allowing researchers to search for papers where a term
appears in sentences with a specific CoreSC label (e.g only in {\em Method}
senetences). This can be used to greatly improve both the precision with which
researchers are able to perform searches for scientific literature and the
accuracy of those searches in terms of relevance to the reader.

The type of a paper ({\em review}, {\em case study}, {\em research}, {\em
perspective}, etc) is another useful feature through which a user can narrow
down the results of a search.  The type of a paper can then be used to augment
queries.  For example, a user may search for a {\em review} paper containing
the keywords ``PCR microfluidics'', or a {\em research} paper with a {\em
hypothesis} containing the keywords ``cerevisiae'' and ``glucose''.   Such
paper types are not yet standardised by journals.  We expect the structure of a
paper to reflect its paper type.  For example, review papers would be expected
to contain a large amount of background material.  In this article, we describe
the application of machine learning (using random forests) to create predictive
models of a paper's type, using the distribution of CoreSC labels found in the
full text of the paper.

This model of paper type is currently in use in our Partridge system, which has
been created as an intelligent full-text search platform for scientific papers.
Partridge (which currently holds XXX papers and is constantly expanding) 
makes use of automatically derived CoreSC sentence labels and
automatically derived paper types, to allow deeper information queries.  We
discuss the reliability of this model of paper types and the insights that have
been gained for the authorship of papers.

\section{Methods}
\label{sec:2}
%how types were chosen,

\subsection*{Collection of scientific articles} Partridge allows users to
upload any paper which is free of copyright restrictions.  For the purpose of
this study we needed a large set of papers that we could label with CoreSC to
investigate how this information assists classification into paper type.  Open
Access (OA) journals provide free read access to their papers, but many do not
permit the user of articles for data mining purposes. The PLoS
journals\footnote{PLoS: \url{http://www.plos.org/}} contain large volumes of OA
literature under a permissive license that allows data mining.  They also use
the Pubmed Central markup schema, which is compatible with SAPIENTA, for papers
published through their journal. The PLoS journals advanced search offers
approximately 50 types of paper through which to restrict the search.  Many of
these paper type categories contain too few papers to be useful, others are too
specific (e.g.``Message from PLoS'') or appear rather adhoc. Indeed it is not
clear how these paper types have been identified. We chose to look at a range
of types, some of which we expect to overlap or have an unclear distinction.
These are namely: Essay, Correspondence, Case Study, Perspective,
Viewpoint, Opinion, Review, Research.

% in conclusions add how this can give recommendations for Plos and other journals 
The papers were downloaded from PLoS using {\em plosget.py}, a short
python script that uses the PLoS RESTful search API. The number of
papers downloaded per paper type category was as follows: 200 Essay,
99 Correspondence, 107 Case Study, 200 Perspective, 74 Viewpoint, 93
Opinion, 312 Review, 200 Research.  These formed a corpus of 1285 %1459-174 after removing synopsis. Should double check the numbers add up
papers.

Figure \ref{fig:coresc_pies} shows the CoreSC content of a review paper and a
research paper randomly selected from the corpus.  The review papers tend to be
made up almost entirely from Background CoreSC sentences.  However, research
papers are much more evenly spread, made up of several different types of
CoreSC.  This investigation suggested that there is almost certainly a
discriminative relationship between CoreSC categories and a paper's type.

\subsection*{Paper Processing and classification}
%Discuss in more detail how Partridge processes papers (PDFX, sentence splitting) 
In order to obtain automatic labeling of sentences in a paper with
CoreSC concepts we first needed to convert the papers to a format that
the the CoreSC classifier, SAPIENTA can analyse.  Currently, SAPIENTA
supports the SciXML and PubMed Central DTDs.  Papers in PDF format are
first converted to XML using PDFX %\cite{PDFX} - no paper known yet
, a free service hosted by the University of Manchester
(\url{http://pdfx.cs.man.ac.uk/}). Once the paper has been split into
sentences, SAPIENTA is used to annotate each sentence with a relevant
CoreSC label. These labels are assigned using a conditional random
fields model based upon a number of features such as the sentence's
location within the paper and pairs and triplets of words found
consecutively within the sentence \cite{Liakata2012}.  The SAPIENTA
system has been trained on a corpus of 265 chemistry and biochemistry
papers and we have done no domain adaptation for the papers in PLoS.

\begin{figure}[t]

\centering

\includegraphics[width=0.75\textwidth]{figures/review_corescs.png}

\includegraphics[width=0.75\textwidth]{figures/report_corescs.png}

\caption{The CoreSC content of a review paper and a research paper randomly selected from the corpus}

\label{fig:coresc_pies} 
\end{figure}

%how the machine learning is conducted (random forest using orange, features, validation methods).

Machine learning to build a classifer of paper type was conducted using random
forest learning \cite{Breiman2001}.  A random forest learning algorithm was
chosen because this is a fast, accurate and widely accepted learning
technology, and because the underlying decision trees can be used to describe
and inspect some of the reasoning behind the predictions made by the final
paper type classifiers.  The feature set consisted of the percentage
composition of the CoreSC labels given to the sentences in each paper i.e. (the
percentage of {\em Background} sentences, the percentage of {\em Hypothesis}
sentences, etc.).  Thus there are 11 features corresponding to the 11 CoreSC
labels.  The random forest learning was conducted using the Orange data mining
library for Python \cite{Curk2005}.  The parameters for the random forest
learner were as the defaults, except with min\_subsets=5 and same\_majority\_pruning=true.  
We used 10-fold cross validation to estimate the precision,
recall and F-measure.

\section{Results and Discussion}
\label{sec:3}

We report the results of the random forest learning as recall, precision and F-measure for each paper type, averaged over a 10-fold cross validation in Table \ref{tab:recallPrecision}. Paper types {\em Research}, {\em Review} and {\em Correspondence} are the most accurate classes, and {\em Viewpoint} is the most difficult paper type to predict.

\begin{table}
\caption{Per-class recall, precision and F-measure micro-averaged over 10-fold cross validation, reporting the results on the held-out validation segment}
\label{tab:recallPrecision}       % Give a unique label
\begin{tabular}{p{2cm}p{2.4cm}p{2cm}p{4.9cm}}
\hline\noalign{\smallskip}
Classes & Recall & Precision & F-measure  \\
\noalign{\smallskip}\svhline\noalign{\smallskip}
Case Study   &     0.243    &    0.245     &   0.244 \\
Correspondence  &     {\bf  0.545}    &    {\bf 0.505}    &    {\bf 0.524} \\ 
Essay     &   0.450     &   0.462    &    0.456 \\
Opinion   &     0.247    &    0.235  &      0.241 \\   
Perspective &       0.255  &      0.304  &       0.277 \\
Research     &   {\bf 0.700}    &    {\bf 0.619}    &    {\bf 0.657} \\
Review      &  {\bf 0.631}     &   {\bf 0.625}     &   {\bf 0.628} \\
Viewpoint   &     0.149     &   0.157  &      0.153 \\
\noalign{\smallskip}\hline\noalign{\smallskip}
\end{tabular}
\end{table}

A confusion matrix for the paper types in given in Table
\ref{tab:confusion}. Research papers are usually predicted to belong
to the Research class, and are sometimes confused with Case
Study. However, Case Study papers are often predicted to be Research
or Review. The ratio between Case study, Research and Review papers is
~1:2:3 so the outcome is not entirely surprising in terms of the data
size effect. We expected Research and Case Study to share similar
paper structures but perhaps the fact that Case Study is equally
confused with Research and Review suggests two distinct types of Case
Study.  We expected the classes Essay, Opinion, Perspective and
Viewpoint to be confused, as the four labels all indicate a paper
containing an author's personal thoughts on an issue, rather than
experiment-driven science. Opinion is confused with Perspective and
Viewpoint. Perspective is confused with almost all classes, except
Research. Viewpoint is a small class and mostly confused with Opinion
and Perspective. From this it would seem that the paper types
\{Opinion, Perspective, Viewpoint\} are very similar and should be
grouped together into one category. Indeed when we trained the model
on a super class consisting of the previous three categories,
performance was improved overall with the following f-measures:
OpinionSuper: 59, Research: 67.6, Review: 63.7, Essay: 44.9,
Correspondence: 46 abd Case Study 24.8.

Essay appears to be distinct enought to justify a class of its own.
Essay is confused with Perspective and Review. Review is the second
best category but is at times confused with Case Study and Essay. A
possible explanation is that Review should cover a wide range of
background material (which we may also expect to find in an Essay),
and could make recommendations based on the material presented (which
we might also expect to find in a Case Study).

\begin{table}
\caption{Confusion matrix summed over 10-fold cross validation, reporting the results on the held-out validation segment. Rows represent true classes, columns represent predicted classes.}
\label{tab:confusion}       % Give a unique label
\begin{tabular}{lllllllll}
\hline\noalign{\smallskip}
      &  Case Study   &     Corresp.      &   Essay  & Opinion &     Perspective    &     Research    &     Review  & Viewpoint \\
\noalign{\smallskip}\svhline\noalign{\smallskip}
Case Study    &    26   &     2     &   11    &    2  &      5 & 26       & 25 &       10 \\
Correspondence  &      4  &      54   &     3    &    9     &   15 & 8   &     4 &       2 \\
Essay      &  8     &   5    &   90    &    2 &       39      &  5 & 46    &    5 \\
Opinion    &    4   &     7 &       2   &     23      &  27     &   10   &  3  &      17 \\
Perspective &       11  &      16    &    49       & 30      &  51 & 6      &  26    &   11 \\
Research      &  19    &    6 &       2 &       8    &    5      &  140  &   12    &    8\\
Review       & 27     &   8     &   35    &    5      &  13    &    21   &  197 &       6\\
Viewpoint    &    7  &      9   &     3     &   19     &   13 & 10     &   2  &      11 \\
\noalign{\smallskip}\hline\noalign{\smallskip}
\end{tabular}
\end{table}

%Discuss why it's possible to predict type from CoreSC (which CoreSC define a class).

We inspected the detail of a single decision tree, grown on the entire dataset, to give further insight into which CoreSC class decisions were responsible for the paper type predictions. This was a very large tree, with a depth of 37 nodes in places.  The first decision was regarding the number of Background sentences in the article. A low Background percentage, of less than 0.694 indicates a Correspondence paper. 28 out of the 31 papers that were classified by this decision as Correpondence were correctly labelled.

The next decision, for higher amounts of Background, was based on the
percentage of Experiment sentences in the paper. For a very low
percentage of Experiment sentences ($<$ 0.061), the papers then
branched into a long side chain of detailed classification decisions,
to separate mostly the Opinion, Viewpoint and Perspective papers from
other Correspondence papers, and a few examples of the remaining
categories. For a higher percentage of Experiment sentences, Research
papers were classified as those that had Observations $>$ 5.6\%, or
Conclusions $<=$ 1.4\%, whereas Case Studies had fewer Observations
and more Conclusions.

The largest node classifiying Essay did so via a route after the low
Experiment decision that asked for a Background $>$ 48\%, but then low
values for Goal, Hypothesis, Result, Observation, Model, Conclusion
and Object. These decisions seem reasonable and agreed with our
expectations of the content of an Essay.

\section{Conclusion}
\label{sec:4}

To summarise, we have demonstrated that paper type can largely be
predicted from the distribution of the automatically obtained CoreSC
sentence labels in the full text of a paper. We have described some of
the particular CoreSC features that determine a paper type
(e.g. Corresponsence characterised by very little Background, Opinions
characterised by low percentage of Experiment) and discussed which of
the paper types are not easily separable in this way. We recommended
for example merging the Opinion, Viewpoint and Perspective articles
into a single category, which increased overall
classification performance. It is also potentially useful to distinguish between two types of Case study.
Analysis of an example tree shows the decision making
process to be complex, but agrees with our general understanding of
paper types. The Partridge online system now allows refinement of
paper search using CoreSC and paper type, both of which can be
intelligently determined using machine learning methods.

%Discuss the potential for intelligent paper search opportunities in general and the future for Partridge.

The potential for automatic extraction of useful features from
scientific papers to assist researchers in their knowledge queries is
now an exciting area for research. Open Access journals that permit
full text mining lead the way in allowing this research to expand and
flourish. In future work we aim to develop and implement a range of
further useful properties that will uncover more of the information
that is hidden within the text of articles, and to use Partridge as a
working engine to demonstrate their usefulness in practice.

\begin{acknowledgement}
Dr Liakata is in receipt of a Leverhulme trust Early Career Fellowship.
\end{acknowledgement}

\bibliographystyle{plain}
\bibliography{partridge}

%\input{referenc}
\end{document}